Tor's Usability for Censorship Circumvention

### 2015-10-06 paper/circumvention-experiment.tex

\begin{abstract}
Tor has grown beyond its original purpose as an anonymity tool and has 
since become an important censorship circumvention tool. It now listed 
as one of the ways that normal people use Tor \cite{whotor}.
We specifically examine its usability as a censorship circumvention tool,
an essential facet for adoption and use.  
We focus our analysis on the connection configuration interface of Tor browser,
as censorship circumvention requires correct transport configurations.
We will conduct a large-scale user study examining 60-100 of users 
on how they navigate Tor's configuration wizard to complete seven browsing tasks 
in three different adversarial settings. Our study combines quantitative measurements (interface
paths taken to success, time to success, and which configuration was chosen) and
qualitative measurements (if users were comfortable with use, what was most confusing, and 
if they would use the browser again). The first phase of the study will evaluate if and how 
easily users can circumvent censorship using Tor Browser. The second phase
of the study will test improvements to the interface and an alternate interface. 
Our goal is to integrate positive usability changes into the Tor Browser. Since
the configuration interface is modular and does not require changes to the Tor Browser
functionality, these changes will be easy to deploy.

@@

\section{Tor's Usability}
As the most widely used anonymity tool today, Tor is primarily known as an anonymity tool. 
For this reason, user studies on Tor have been solely about, the usability of Tor as an 
anonymity tool. Norcie~\cite{norcie2012eliminating} conducted an experiment which identified 
``stopping points'' in Tor browser,  documenting points when people would get frustrated enough 
to stop using Tor. To our knowledge, that has been the only published user study of Tor. 
Since then, Tor has had a lot of updates.There have been been no published usability evaluations of
Tor Browser since the introduction of the 4.0 series, which introduced radical UI changes. 
Lee and Fifield \cite {uxsprint} ran a small pilot study testing the download, install, and user interface of Tor Browser. 
This study uncovered a number of bugs and stopping points. Changes made are reflected in series 
5.1 and later.

There are still many features of Tor that are left unevaluated through user research---such as advanced web tasks (such as accessing hidden services), the configuration menu to connect to the Tor network, configuring for automatic updates, and identity/cookie management. Rather than selecting the features to study in isolation, 
we decided to focus on an important use case of Tor browser, censorship circumvention. 
Internet censorship is shown to be pervasive across the world today \cite{faris2008measuring}, with 
some countries blocking censorship circumvention tools as well \cite{winter2012great}. 
To our knowledge, this is the first user study investigating the usability of Tor as a 
censorship circumvention tool.

@@

Our study will test this hypotheses by measuring if the connection
interface successfully guides users to correctly configuring their browser.\\

@@

\noindent {\bfseries Experiment Logistics} 
The IRB protocol to run this user study has been approved (2014-12-6995). 
We plan to recruit from 100-200 users for the purpose of this study, 
making it the largest user study of Tor to date.  This study will be conducted at the
Experimental Social Science Laboratory (Xlab)
at the University of California, Berkeley, which consists of 36 laptops,
separated by cubicle walls. We will use individual host firewalls to simulate
censorship environments and will record computer screens to capture 
user activity. The total length of the experiment, including briefing, completing the censorship 
circumvention tasks, exit survey, and debriefing, will be about an hour.
Participants will be compensated \$30 for their time, which covers
minimum wage for an hour and any transportation costs to the lab.  \\

\noindent {\bfseries Experiment Flow} 
We will run three sets of the experiment, the first set to test the current configuration
interface, a second set to test the improved interface, and a third set to test the 
alternate configuration interface. The experiment flow will be the same across all 
sets of experiments, but the configuration interface will be different across sets.

The experiment begins with the participants being informed that they are in a
simulated censorship
environment, where some --- but not all --- websites are blocked. We will
instruct them to visit a non-blocked website and a blocked website on a
``standard'' browser (one that is not used for censorship circumvention, such
as Chrome, Safari, or Firefox) to illustrate the situation.
Then, the participants will find out the role they they will be playing in the study.
This role is that of a regular citizen, who is trying to visit blocked websites
for non-critical leisurely purposes,
and that there is a minimal chance of risk. This information is given
to achieve
consistency across participants: behaviors may change if the tasks were 
critical or if there were great risks associated with configuration mistakes.

@@

\noindent {\bfseries Censorship Environment Simulation} 
We plan to simulate three censorship environments.
They are informed by our experience with pluggable transports
and knowledge of commonly seen censorship techniques.
They are not meant perfectly to replicate the network environment
in any particular country. Although inspired by reality, these
abstract simulations are intended to require distinct configurations
of the Tor Browser from our participants.

@@

The tasks were initially inspired by the top Alexa sites, 
an indication of representative and relevant browsing behavior. 
From these, we selected sites which were commonly censored, 
but filtered tasks that would require a participant to reveal private information 
(such as login information) for ethical considerations. These tasks
were further refined after performing a pilot study of the experiment. 
We hope the tasks convey an example of how of a user might 
browse the Internet in a censored environment.

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item Google search for the population of Zimbabwe. 
\item On YouTube, find a video playing Bach's ``Ode to Joy.''
\item Find the Amazon best-sellers in ``Movies \& TV.''
\item On Yahoo, find the exchange rate of Dollars to Euros.
\item Find the Wikipedia ``History'' portal's featured article. 
\item On Twitter, find the currently trending topics.
\item On Bing Maps, find directions from Times Square to Carnegie Hall.\\
\end{itemize}

\noindent {\bfseries Improved and Alternative Interfaces} We have yet to design the 
improvements to the interface, as they will be based on the data collected from 
the user studies on the current configuration interface. In general, we will measure the 
time it took to configure correctly, and help users avoid common mistakes. An alternate 
interface to automate the configuration process for the user has been proposed in the 
community. There issue is that the automation make the connection process less transparent, 
and since there are countries who log internet activity, automated attempts to obviously
use the Tor network could have the potential to be incriminating.

@@

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item {\bfseries Test the Tor configuration interface} We will perform the largest user study 
of Tor, measuring Tor configuration in three different adversarial settings. 
\item {\bfseries Test an improved configuration interface} With the measurements collected
from testing the interface as-is, we will be design and test improvements to minimize
time taken, paths taken, and error states reached.
\item {\bfseries Test an alternative configuration interface} A Tor developer has offered
to design a mock-up interface for an alternative configuration interface which will
greatly automate the process. There are tradeoffs between ease of use and transparency
of the systems, and ethical considerations with loggable errors resulting from automated configuration. 
\item {\bfseries Push changes} We have the support of Tor developers
to improve this interface. Additionally, since the configuration interface does not require 
any changes to the Tor Browser functionality, improvements to the interface will
be easy to deploy. 
\end{itemize}

>>>>

\begin{abstract}
Tor has grown beyond its original purpose as an anonymity tool and has 
since become an important censorship circumvention tool. It is now listed
as one of the ways that normal people use Tor \cite{whotor}.
We specifically examine its usability as a censorship circumvention tool,
an essential facet for adoption and use.  
We focus our analysis on the connection configuration interface of Tor browser,
as censorship circumvention requires correct transport configurations.
We will conduct a large-scale user study examining 60-100 of users 
on how they navigate Tor's configuration wizard to complete seven browsing tasks 
in three different adversarial settings. Our study combines quantitative measurements (interface
paths taken to success, time to success, and which configuration was chosen) and
qualitative measurements (if users were comfortable with use, what was most confusing, and 
if they would use the browser again). The first phase of the study will evaluate if and how 
easily users can circumvent censorship using Tor Browser. The second phase
of the study will test improvements to the interface and an alternate interface. 
Our goal is to integrate positive usability changes into the Tor Browser. Since
the configuration interface is modular and does not require changes to the Tor Browser
functionality, these changes will be easy to deploy.

@@

\section{Tor's Usability}
As the most widely used anonymity tool today, Tor is primarily known for this aspect of its functionality.
For this reason, user studies on Tor have been solely about the usability of Tor as an
anonymity tool. Norcie~\cite{norcie2012eliminating} conducted an experiment which identified 
``stopping points'' in Tor browser,  documenting points when people would get frustrated enough 
to stop using Tor. To our knowledge, that has been the only published user study of Tor. 
Since then, Tor has had a lot of updates. There have been been no published usability evaluations of
Tor Browser since the introduction of the 4.0 series, which introduced radical UI changes. 
Lee and Fifield \cite {uxsprint} ran a small pilot study testing the download, install, and user interface of Tor Browser. 
This study uncovered a number of bugs and stopping points. Changes made are reflected in series 
5.1 and later.

There are still many features of Tor that are left unevaluated through user
research --- such as advanced web tasks (e.g., accessing hidden services), the
configuration menu to connect to the Tor network, configuring automatic
updates, and identity/cookie management. Rather than selecting the features to
study in isolation,
we decided to focus on an important use case of Tor browser: censorship circumvention. 
Internet censorship is pervasive across the world today \cite{faris2008measuring}, with 
some countries blocking censorship circumvention tools as well \cite{winter2012great}. 
To our knowledge, this is the first user study investigating the usability of Tor as a 
censorship circumvention tool.

@@

Our study will test this hypothesis by measuring if the connection
interface successfully guides users to correctly configuring their browser.\\

@@

\noindent {\bfseries Experiment Logistics} 
The IRB protocol to run this user study has been approved (2014-12-6995). 
We plan to recruit 100-200 users for the purpose of this study,
making it the largest user study of Tor to date.  This study will be conducted at the
Experimental Social Science Laboratory (Xlab)
at the University of California, Berkeley, which consists of 36 laptops,
separated by cubicle walls. We will use individual host firewalls to simulate
censorship environments and will record computer screens to capture 
user activity. The total length of the experiment, including briefing, completing the censorship 
circumvention tasks, exit survey, and debriefing, will be about an hour.
Participants will be compensated \$30 for their time, which covers
minimum wage for an hour and any transportation costs to the lab.  \\

\noindent {\bfseries Experiment Flow} 
We will conduct three separate experiments as part of this study:
the first set to test the current configuration
interface, a second set to test the improved interface, and a third set to test the 
alternate configuration interface. The experiment flow will be the same across all 
sets of experiments, but the configuration interface will be different across sets.

The experiment begins with the participants being informed that they are in a
simulated censorship
environment, where some --- but not all --- websites are blocked. We will
instruct them to visit a non-blocked website and a blocked website on a
``standard'' browser (one that is not used for censorship circumvention, such
as Chrome, Safari, or Firefox) to illustrate the situation.
Then, the participants will find out the role they they will be playing in the study.
This role is that of a regular citizen who is trying to visit blocked websites
for non-critical leisurely purposes,
and for whom the chance of risk is minimal. This information is given
to achieve
consistency across participants: behaviors may change if the tasks were 
critical or if there were great risks associated with configuration mistakes.

@@

\noindent {\bfseries Censorship Environment Simulation} 
We plan to simulate three censorship environments.
They are informed by our experience with pluggable transports
and knowledge of commonly seen censorship techniques.
They are not meant to perfectly replicate the network environment
in any particular country. Although inspired by reality, these
abstract simulations are intended to require distinct configurations
of the Tor Browser from our participants.

@@

The tasks were inspired by the top Alexa sites,
an indication of representative and relevant browsing behavior. 
From these, we selected sites which were commonly censored, 
but filtered tasks that would require a participant to reveal private information 
(such as login information) for ethical considerations. These tasks
were further refined after performing a pilot study of the experiment. 
We hope the tasks convey an example of how of a user might 
browse the Internet in a censored environment.

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item Google search for the population of Zimbabwe. 
\item On YouTube, find a video playing Bach's ``Ode to Joy.''
\item Find the Amazon best-sellers in ``Movies \& TV.''
\item On Yahoo, find the exchange rate of dollars to euros.
\item Find the Wikipedia ``History'' portal's featured article. 
\item On Twitter, find the currently trending topics.
\item On Bing Maps, find directions from Times Square to Carnegie Hall.\\
\end{itemize}

\noindent {\bfseries Improved and Alternative Interfaces} We have yet to design the 
improvements to the interface, as they will be based on the data collected from 
the user studies on the current configuration interface. In general, we will measure the 
time it took to configure correctly, and help users avoid common mistakes. An alternate 
interface to automate the configuration process for the user has been proposed in the 
community. One concern about it is that the automation makes the connection process less transparent, 
and since there are countries who log internet activity, automated attempts to obviously
use the Tor network could have the potential to be incriminating.

@@

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item {\bfseries Test the Tor configuration interface} We will perform the largest user study 
of Tor, measuring Tor configuration in three different adversarial settings. 
\item {\bfseries Test an improved configuration interface} With the measurements collected
from testing the interface as-is, we will design and test improvements to minimize
time taken, paths taken, and error states reached.
\item {\bfseries Test an alternative configuration interface} A Tor developer has offered
to design a mock-up interface for an alternative configuration interface which will
greatly automate the process. There are tradeoffs between ease of use and transparency
of the systems, and ethical considerations with loggable errors resulting from automated configuration. 
\item {\bfseries Push changes} We have the support of Tor developers
to improve this interface. Additionally, since the configuration interface does not require 
any changes to the Tor Browser functionality, improvements to the interface will
be easy to deploy. 
\end{itemize}
